Spectroscopic confirmation of very-wide binaries and large-separation comoving stars from \textsl{Gaia DR1}

### 2017-07-12 paper/comoving-followup.tex
discussion

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information supports the belief that the two stars are
truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\ (as computed in \citealt{Oh:2017})m and the log-FML-ratio including the radial velocity data as \llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$---128 of
the 311 comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
observed pairs.
In detail, for visualization, we sample from the error distributions over proper
motion, parallax, and radial velocity for each star individually and compute the
magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
samples.
We plot the median of the resulting distribution of velocity-differences as
dark markers (circles).
The error bars show the 15th and 85th percentiles of the distributions for a few
randomly chosen pairs.
Note that pairs with large 3D velocity differences have large (negative) changes
to their log-likelihood-ratio values.

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{genuine_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
    the density of all \tgas\ sources within 200 pc; similar to
    \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
    pairs.
    Several interesting pairs that include main-sequence and giant component
    stars are highlighted in the text
    (\sectionname~\ref{sec:interesting-pairs}).
    \label{fig:genuine-cmd}}
\end{figure}

@@

We naively expect that comoving, coeval stars should lie along lines parallel to
the main sequence in color-magnitude space. The majority of our star pairs
appear to follow this expectation, with the exception of a few visible outliers.
For example, there appear to be at least a few main-sequence--giant-branch
pairs, and a few pairs that appear to cross the main-sequence.
We highlight a few of these pairs below.

\subsubsection{HD 95088 \& HD 92278}

\todo{Highlight the G subgiant-A pair}

\subsubsection{HD 151090 \& HIP 81988}

\todo{G-subgiant - K-dwarf - well known high proper-motion comoving pair}

\subsubsection{HD 54318 \& TYC 3792-658-1}

\todo{Highlight the largest mag difference MS pair - K-dwarf - B-star}

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist-rave.pdf}
  \end{center}
  \caption{%
    Same as \figurename~\ref{fig:separation}, but for our sample of genuine
    comoving stars combined with all genuine comoving stars identified from the
    RAVE overlap.
    \label{fig:separation-RAVE}}
\end{figure}

% \begin{figure}[htbp]
%   \begin{center}
%     \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
%   \end{center}
%   \caption{%
%     Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
%     the density of all \tgas\ sources within 200 pc; similar to
%     \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
%     pairs.
%     Several interesting pairs that include main-sequence and giant component
%     stars are highlighted in the text
%     (\sectionname~\ref{sec:interesting-pairs}).
%     \label{fig:genuine-cmd}}
% \end{figure}

@@

\todo{Do I want to do isochrone fitting for all of them?}

@@

\acknowledgements
John Brewer (Yale),
Jason Curtis (Columbia),
James Davenport (WWU) for PyDIS,
David W. Hogg (NYU),
Melissa Ness (MPIA)

>>>>

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information supports the belief that the two stars are
truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\ (as computed in \citealt{Oh:2017})m and the log-FML-ratio including the radial velocity data as \llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$ ---
\ncomoving of the \npairsobs\ comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
observed pairs.
In detail, for visualization, we sample from the error distributions over proper
motion, parallax, and radial velocity for each star individually and compute the
magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
samples.
We plot the median of the resulting distribution of velocity-differences as
dark markers (circles).
The error bars show the 15th and 85th percentiles of the distributions for a few
randomly chosen pairs.
Note that pairs with large 3D velocity differences have large (negative) changes
to their log-likelihood-ratio values.

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{genuine_cmd.pdf}
  \end{center}
  \caption{%
    Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
    the density of all \tgas\ sources within 200 pc; similar to
    \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
    pairs.
    Several interesting pairs that include main-sequence and giant component
    stars are highlighted in the text
    (\sectionname~\ref{sec:interesting-pairs} and
    \figurename~\ref{fig:highlighted-cmd}).
    \label{fig:genuine-cmd}}
\end{figure}

@@

We naively expect that comoving, coeval stars should lie along lines parallel to
the main sequence in color-magnitude space. The majority of our star pairs
appear to follow this expectation, with the exception of a few visible outliers.
For example, there appear to be at least a few main-sequence--giant-branch
pairs, and a few pairs that appear to cross the main-sequence.
We highlight a few of these pairs below and in
\figurename~\ref{fig:highlighted-cmd}.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{highlighted_cmd.pdf}
  \end{center}
  \caption{%
    Same as \figurename~\ref{fig:genuine-cmd}, but only three highlighted
    comoving star pairs are shown.
    \label{fig:highlighted-cmd}}
\end{figure}

\subsubsection{HD 95088 -- HD 92278}

HD 95088 HD 92278
Angular sep.: 4.49 deg
Chord length: 8.66 pc
3D sep.: 9.53 pc
parallaxes: 9.00 +/- 0.50, 9.32 +/- 0.51

This pair is separated by $8.7~\pc$ in chord length ($4.5^\circ$ on the sky) and
have consistent distances within their uncertainties of $\approx 110~\pc$.
From visual classification of the stars' location in the color--magnitude
diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
comprised of an A-type dwarf and a slightly less massive G-type subgiant.
If this is indeed the case, the two stars could therefore not be coeval, as the
more massive A star has not begun its post-main-sequence evolution.
One possible explanation is that this pair is a false-positive in three
dimensions, in which case they are unassociated but just appear to be comoving
with small relative velocity at present day.

Other possibilities stem from the possibility that the comoving pair is the
outcome of an interaction between more stars.
For example, an interaction between two wide binaries could result in an
exchange of members, a hardening of one resulting binary, and a widening of the
other binary (see, e.g., \citealt{Nathan??}).
Or, the initial system could have been an unstable three-body system that
resulted in a tight, inner binary and a wide outer binary;
if the inner binary were close enough, it could have merged and formed a blue
straggler star (e.g., \citealt{??}), leaving the wide binary untouched.

These explanations are all speculative at the moment; high-resolution
spectroscopy of these stars --- and precise measurements of the stellar
parameters and chemical abundances --- may help to distinguish these
possibilities.

\subsubsection{HD 79247 -- TYC 3807-1439-1}

This pair is separated by $XX~\pc$ in chord length ($XX^\circ$ on the sky) and
have consistent distances within their uncertainties of $\approx XX~\pc$.
From visual classification of the stars' location in the color--magnitude
diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
comprised of a G-type subgiant and a G-type dwarf.

\subsubsection{HD 54318 -- TYC 3792-658-1}

TYC 3792-658-1 HD 54318
Angular sep.: 2.51 deg
Chord length: 8.45 pc
3D sep.: 8.73 pc
parallaxes: 5.20 +/- 0.23, 5.30 +/- 0.32

This pair is separated by $8.5~\pc$ in chord length ($2.5^\circ$ on the sky) and
have consistent distances within their uncertainties of $\approx 190~\pc$.

From visual classification of the stars' location in the color--magnitude
diagram (\figurename~\ref{fig:highlighted-cmd}), the pair appears to be
comprised of a late-A-/early-B-type dwarf and a G-type dwarf.

@@

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-with-rave.pdf}
  \end{center}
  \caption{%
    \label{fig:separation-RAVE}}
\end{figure}

Caution about selection criteria, false-positives at large separation...

@@

@@

\acknowledgements

Dave Zurek, Keith, Ruth, Kelle
John Brewer (Yale),
Jason Curtis (Columbia),
James Davenport (WWU) for PyDIS,
David W. Hogg (NYU),
Melissa Ness (MPIA)
